Birth Spacing in the Presence of Son Preference and Sex-Selective \\ Abortions: India's Experience over Four Decades

### 2018-04-19 presentations/2018-paa-poster.tex
Shorten background block

Son preference is often associated with shorter spacing after the 
births of girls than boys, shorter spacing is, in turn, associated 
with worse health outcomes for girls and mothers.

However, the introduction of prenatal sex determination fundamentally changed the 
relationship between son preference and birth spacing.
Couples, who had shorter birth spacing before because of son preference,
now have access to prenatal sex determination and sex-selective abortions,
and each abortion increases birth spacing by six months to a year.

As a result, we now have a situation where we may observer \emph{longer} spacing
for families with daughters than for families with sons, precisely because
strong son preference leads them to use sex selection.

To further complicate matters, we can still observe shorter birth spacing 
after the births of daughters as a representation of son preference for 
families who---for one reason or another---do not use prenatal sex selection.

@@

\item[Main idea:]  in the absence of sex selection son preference leads to shorter 
spacing after the birth of a girl than after the birth of a boy, whereas 
son preference increases the spacing after the birth of a girl relative 
to after the birth of a boy when prenatal sex selection is available.

@@

\item[Central question:] how spacing patterns changed with access
to prenatal sex determination became available.
Specifically, did the spacing after girls increase relative to the spacing
after boys?
\end{description}

>>>>

Son preference has been associated with shorter spacing after the 
births of girls than boys.

\vskip3ex

Access to prenatal sex determination fundamentally changed the 
relationship between son preference and birth spacing because each abortion 
increases birth spacing by six months to a year.

\vskip3ex

We may therefore now observe \alert{longer} spacing after daughters than sons, 
precisely because strong son preference leads to the use of sex selection.

\vskip3ex

To complicate matters, shorter birth spacing after daughters may still 
be a representation of son preference for 
families who do not use prenatal sex selection.

@@

\item[Main idea:]  in the absence of sex selection son preference leads to shorter 
spacing after the birth of a girl than after the birth of a boy, whereas 
son preference increases the spacing after the birth of a girl relative 
to after a boy when prenatal sex selection is available.

@@

\item[Central question:] how did spacing patterns changed with access
to prenatal sex determination became available?
Specifically, did the spacing after girls increase relative to the spacing
after boys?
\end{description}
